---
title: "HW2"
author: "Steven Leone"
format: html
editor: visual
---


This is the Github Repository for STATS 506 HW2: https://github.com/leonestv-byte/5å06_hw2


## Exercise 1: Random Walk with Momentum


A standard random walk assumes that each step is independent of all previous steps. In many real processes, however, the current movement may depend on previous movement.

For example, a recent study of lab-grown diamond substrates uses a two-dimensional random walk with momentum to model the tendency of stress orientations in consecutive layers of a diamond crystal to remain similar. Higher momentum corresponds to stronger persistence between consecutive layers.

At each step, the particle moves one unit in one of four directions:

* right: $(+1, 0)$
* left: $(-1, 0)$
* up: $(0, +1)$
* down: $(0, -1)$

For the first step, choose one of the four directions with equal probability (which means, 1/4). For subsequent steps, with probability p, continue; for 1-p, choose another.

a) 
```{r}
directions <- c("right", "left", "up", "down")

choose_new_direction <- function() {
  random_integer <- sample(1:4, 1)
  return(directions[random_integer])
}

step <- function(direction, coordinate_x, coordinate_y) {
  delta <- switch(direction,
    "right" = c(1, 0),
    "left"  = c(-1, 0),
    "up"    = c(0, 1),
    "down"  = c(0, -1),
    stop("direction must be 'right', 'left', 'up', or 'down'")
  )

  coordinate_x <- coordinate_x + delta[1]
  coordinate_y <- coordinate_y + delta[2]
  
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}

momentum_walk <- function(n, p) {
  
  if (n < 1) {
    stop("n must be greater than 1.")
  }
  if (p < 0 || p > 1) {
    stop("p must be in between 0 and 1.")
  }
  
  coordinate_x <- 0
  coordinate_y <- 0
  
  # First Step
  direction <- choose_new_direction()
  result <- step(direction, coordinate_x, coordinate_y)
  coordinate_x <- result$coordinate_x
  coordinate_y <- result$coordinate_y
  
  # Subsequent Steps
  for (i in 2:n) {
    my_bool <- as.logical(rbinom(1, size = 1, prob = p))
    if (my_bool) {
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    } else {
      direction <- choose_new_direction()
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    }
  }
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}
```

b) Implement the simulation.

```{r}
directions <- c("right", "left", "up", "down")

choose_new_direction <- function() {
  random_integer <- sample(1:4, 1)
  return(directions[random_integer])
}

step <- function(direction, coordinate_x, coordinate_y) {
  delta <- switch(direction,
    "right" = c(1, 0),
    "left"  = c(-1, 0),
    "up"    = c(0, 1),
    "down"  = c(0, -1),
    stop("direction must be 'right', 'left', 'up', or 'down'")
  )

  coordinate_x <- coordinate_x + delta[1]
  coordinate_y <- coordinate_y + delta[2]
  
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}

momentum_walk1 <- function(n, p) {
  
  if (n < 1) {
    stop("n must be greater than 1.")
  }
  if (p < 0 || p > 1) {
    stop("p must be in between 0 and 1.")
  }
  
  coordinate_x <- 0
  coordinate_y <- 0
  
  # First Step
  direction <- choose_new_direction()
  result <- step(direction, coordinate_x, coordinate_y)
  coordinate_x <- result$coordinate_x
  coordinate_y <- result$coordinate_y
  
  # Subsequent Steps
  for (i in 2:n) {
    my_bool <- as.logical(rbinom(1, size = 1, prob = p))
    if (my_bool) {
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    } else {
      direction <- choose_new_direction()
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    }
  }
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}

```

```{r}

momentum_walk2 <- function(n, p) {
  if (n < 1) {
    stop("n must be greater than 1.")
  }
  if (p < 0 || p > 1) {
    stop("p must be in between 0 and 1.")
  }
  
  direction <- choose_new_direction()
  instructions <- sapply(seq_len(n), function(i) {
    if (i > 1 && as.logical(rbinom(1, size = 1, prob = p))) {
      direction <<- choose_new_direction()
    }
    return(direction)
  })
  
  delta_x <- c(right = 1, left = -1, up = 0, down = 0)
  delta_y <- c(right = 0, left = 0, up = 1, down = -1)

  list(
    coordinate_x = sum(delta_x[instructions]),
    coordinate_y = sum(delta_y[instructions])
  )
}
```


c)

```{r}
set.seed(979)
momentum_walk1(10, 0.6)

set.seed(979)
momentum_walk1(10,0.6)
```


## Exercise 2: Monte Carlo Simulation of Campus Bus Arrivals



## Exercise 3: Linear Regression

```{r}
youtube <- read.csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv"
)
```

```{r}
model <- lm(dislike_count ~ category_id + comment_count, data=youtube)
```
