---
title: "HW2"
author: "Steven Leone"
format: html
editor: visual
---

This is the Github Repository for STATS 506 HW2: https://github.com/leonestv-byte/5å06_hw2

## Exercise 1: Random Walk with Momentum

A standard random walk assumes that each step is independent of all previous steps. In many real processes, however, the current movement may depend on previous movement.

For example, a recent study of lab-grown diamond substrates uses a two-dimensional random walk with momentum to model the tendency of stress orientations in consecutive layers of a diamond crystal to remain similar. Higher momentum corresponds to stronger persistence between consecutive layers.

At each step, the particle moves one unit in one of four directions:

- right: $(+1, 0)$
- left: $(-1, 0)$
- up: $(0, +1)$
- down: $(0, -1)$

For the first step, choose one of the four directions with equal probability (which means, 1/4). For subsequent steps, with probability p, continue; for 1-p, choose another.

a)  

```{r}
directions <- c("right", "left", "up", "down")

choose_new_direction <- function() {
  random_integer <- sample(1:4, 1)
  return(directions[random_integer])
}

step <- function(direction, coordinate_x, coordinate_y) {
  delta <- switch(direction,
    "right" = c(1, 0),
    "left"  = c(-1, 0),
    "up"    = c(0, 1),
    "down"  = c(0, -1),
    stop("direction must be 'right', 'left', 'up', or 'down'")
  )

  coordinate_x <- coordinate_x + delta[1]
  coordinate_y <- coordinate_y + delta[2]
  
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}

momentum_walk <- function(n, p) {
  
  if (n < 1) {
    stop("n must be greater than 1.")
  }
  if (p < 0 || p > 1) {
    stop("p must be in between 0 and 1.")
  }
  
  coordinate_x <- 0
  coordinate_y <- 0
  
  # First Step
  direction <- choose_new_direction()
  result <- step(direction, coordinate_x, coordinate_y)
  coordinate_x <- result$coordinate_x
  coordinate_y <- result$coordinate_y
  
  # Subsequent Steps
  for (i in 2:n) {
    my_bool <- as.logical(rbinom(1, size = 1, prob = p))
    if (my_bool) {
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    } else {
      direction <- choose_new_direction()
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    }
  }
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}
```

b)  Implement the simulation.

```{r}
directions <- c("right", "left", "up", "down")

choose_new_direction <- function() {
  random_integer <- sample(1:4, 1)
  return(directions[random_integer])
}

step <- function(direction, coordinate_x, coordinate_y) {
  delta <- switch(direction,
    "right" = c(1, 0),
    "left"  = c(-1, 0),
    "up"    = c(0, 1),
    "down"  = c(0, -1),
    stop("direction must be 'right', 'left', 'up', or 'down'")
  )

  coordinate_x <- coordinate_x + delta[1]
  coordinate_y <- coordinate_y + delta[2]
  
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}

momentum_walk1 <- function(n, p) {
  
  if (n < 1) {
    stop("n must be greater than 1.")
  }
  if (p < 0 || p > 1) {
    stop("p must be in between 0 and 1.")
  }
  
  coordinate_x <- 0
  coordinate_y <- 0
  
  # First Step
  direction <- choose_new_direction()
  result <- step(direction, coordinate_x, coordinate_y)
  coordinate_x <- result$coordinate_x
  coordinate_y <- result$coordinate_y
  
  # Subsequent Steps
  for (i in 2:n) {
    my_bool <- as.logical(rbinom(1, size = 1, prob = p))
    if (my_bool) {
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    } else {
      direction <- choose_new_direction()
      result <- step(direction, coordinate_x, coordinate_y)
      coordinate_x <- result$coordinate_x
      coordinate_y <- result$coordinate_y
    }
  }
  return(list(coordinate_x = coordinate_x, coordinate_y = coordinate_y))
}

```

```{r}

momentum_walk2 <- function(n, p) {
  if (n < 1) {
    stop("n must be greater than 1.")
  }
  if (p < 0 || p > 1) {
    stop("p must be in between 0 and 1.")
  }
  
  direction <- choose_new_direction()
  instructions <- sapply(seq_len(n), function(i) {
    if (i > 1 && as.logical(rbinom(1, size = 1, prob = p))) {
      direction <<- choose_new_direction()
    }
    return(direction)
  })
  
  delta_x <- c(right = 1, left = -1, up = 0, down = 0)
  delta_y <- c(right = 0, left = 0, up = 1, down = -1)

  list(
    coordinate_x = sum(delta_x[instructions]),
    coordinate_y = sum(delta_y[instructions])
  )
}
```

c)  

```{r}

seeds <- sample(1:1000, 100)

for (seed in seeds) {
  n = 10
  p = 0.6
  set.seed(seed)
  result <- momentum_walk1(n, p)
  x_one <- result$coordinate_x
  y_one <- result$coordinate_y
  set.seed(seed)
  result <- momentum_walk1(n, p)
  x_two <- result$coordinate_x
  y_two <- result$coordinate_y
  stopifnot(identical(x_one, x_two))
  stopifnot(identical(y_one, y_two))
}

print("All momentum walk 1 simulations passed and checked for equality! n = 10")

for (seed in seeds) {
  n = 10
  p = 0.6
  set.seed(seed)
  result <- momentum_walk2(n, p)
  x_one <- result$coordinate_x
  y_one <- result$coordinate_y
  set.seed(seed)
  result <- momentum_walk2(n, p)
  x_two <- result$coordinate_x
  y_two <- result$coordinate_y
  stopifnot(identical(x_one, x_two))
  stopifnot(identical(y_one, y_two))
}

print("All momentum walk 2 simulations passed and checked for equality! n = 10")


for (seed in seeds) {
  n = 100
  p = 0.6
  set.seed(seed)
  result <- momentum_walk1(n, p)
  x_one <- result$coordinate_x
  y_one <- result$coordinate_y
  set.seed(seed)
  result <- momentum_walk1(n, p)
  x_two <- result$coordinate_x
  y_two <- result$coordinate_y
  stopifnot(identical(x_one, x_two))
  stopifnot(identical(y_one, y_two))
}

print("All momentum walk 1 simulations passed and checked for equality! n = 100")

for (seed in seeds) {
  n = 100
  p = 0.6
  set.seed(seed)
  result <- momentum_walk2(n, p)
  x_one <- result$coordinate_x
  y_one <- result$coordinate_y
  set.seed(seed)
  result <- momentum_walk2(n, p)
  x_two <- result$coordinate_x
  y_two <- result$coordinate_y
  stopifnot(identical(x_one, x_two))
  stopifnot(identical(y_one, y_two))
}

print("All momentum walk 2 simulations passed and checked for equality! n = 100")
```

d)

```{r}
ps = c(0, 0.5, 0.9)

distances <- c(0, 0, 0)

i <- 1
for (p in ps) {
  result <- momentum_walk1(1000, p)
  d <- sqrt((result$coordinate_x **2) + (result$coordinate_y **2))
  distances[i] <- d
  i <- i + 1
}
print(distances)
```

Increasing momentum increases distance. With p at 0.9, we rarely change direction, and do not retread ground - we see this with a 164.1 distance covered.


e) 
```{r}
ps = c(0, 0.25, 0.5, 0.75,  0.9)

distances <- c(0, 0, 0, 0, 0)

i <- 1
for (p in ps) {
  d <- 0
  for (j in 1:1000) {
    result <- momentum_walk1(1000, p)
    d <- d + sqrt((result$coordinate_x **2) + (result$coordinate_y **2))
  }
  d <- d / 1000
  distances[i] <- d
  i <- i + 1
}
print(distances)

data.frame(ps, distances)
knitr::kable(data.frame(ps, distances))
```
There is more distance as p increases.

f) 

```{r}
install.packages("microbenchmark")
library(microbenchmark)


```
```{r}

n = 1000
p = 0.6
benchmark_results <- microbenchmark(
  v1 = momentum_walk1(n, p),
  v2 = momentum_walk2(n, p)
)
print(benchmark_results)

n = 100000
p = 0.6
benchmark_results <- microbenchmark(
  v1 = momentum_walk1(n, p),
  v2 = momentum_walk2(n, p)
)
print(benchmark_results)

```
Version 1, with the for loop is faster. The difference is larger with larger input, as we rack up more time with more runs. The algorithms have different time complexities, which makes the first one faster.

## Exercise 2: Monte Carlo Simulation of Campus Bus Arrivals


```{r}
rows <- sample((1:2), 10000, replace = TRUE)

simulated_day <- function(i) {
  six_eight <- sum(rpois(n = 3, lambda = 4))
  nine <- sum(rnorm(n = 1, mean = 45, sd = 4))
  ten_three <- sum(rpois(n = 6, lambda = 15))
  four <- sum(rnorm(n = 1, mean = 55, sd = 5))
  five_ten <- sum(rpois(n = 6, lambda = 9))
  return(six_eight + nine + ten_three + four + five_ten)
}

rows <- sapply(rows, simulated_day)
print(mean(rows))
```


## Exercise 3: Linear Regression

```{r}
youtube <- read.csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv"
)
```

```{r}
model <- lm(dislike_count ~ category_id + comment_count, data=youtube)
```
